\thetitle

### 2008-01-21 models.tex
Change Indvidual models to Trace.
Redone intro; added paragraph to related work; etc.
In the process of overhauling the Methodology section.

\multicolumn{1}{c}{\textbf{Model}} &

@@

\multirow{8}{*}[2.5pt]{Node}
& \multirow{1}{*}[-0.05em]{Individual} &

@@

& \multirow{1}{*}[-0.05em]{Nonparametric} &

@@

& \multirow{1}{*}[-0.05em]{Parametric} &

@@

\raisebox{1.5pt}{$\centerdot$} generic flows assigned to nodes according to fitted BiPareto model\raisebox{1pt}{\footnotesize*} \\
\raisebox{1.5pt}{$\centerdot$} session start time assigned according to Poisson arrival model\raisebox{1pt}{\footnotesize*}

@@

& \multirow{1}{*}[-0.05em]{Uniform} &

@@

\multirow{11}{*}[2.5pt]{Flow}
& \multirow{1}{*}[-0.05em]{Individual} &

@@

\raisebox{1.5pt}{$\centerdot$} flow size is sampled from fitted BiPareto model\raisebox{1pt}{\footnotesize*} \\
\raisebox{1.5pt}{$\centerdot$} start time relative to previous flow in session sampled from fitted Lognormal model\raisebox{1pt}{\footnotesize*} \\

@@

\multirow{5}{*}[2.5pt]{Packet}
& \multirow{1}{*}[-0.05em]{Individual} &

@@

\raisebox{1.5pt}{$\centerdot$} packet sizes and intervals drawn from fitted Exponential distributions\raisebox{1pt}{\footnotesize*}

>>>>

\multicolumn{1}{c}{\textbf{Sub-model}} &

@@

\multirow{8}{*}[2.5pt]{\class{Node}}
& \multirow{1}{*}[-0.05em]{\submodel{Trace}} &

@@

& \multirow{1}{*}[-0.05em]{\submodel{Nonparametric}} &

@@

& \multirow{1}{*}[-0.05em]{\submodel{Parametric}} &

@@

\raisebox{1.5pt}{$\centerdot$} generic flows assigned to nodes according to fitted BiPareto model\raisebox{1.5pt}{\scriptsize*} \\
\raisebox{1.5pt}{$\centerdot$} session start time assigned according to Poisson arrival model\raisebox{1.5pt}{\scriptsize*}

@@

& \multirow{1}{*}[-0.05em]{\submodel{Uniform}} &

@@

\multirow{11}{*}[2.5pt]{\class{Flow}}
& \multirow{1}{*}[-0.05em]{\submodel{Trace}} &

@@

\raisebox{1.5pt}{$\centerdot$} flow size is sampled from fitted BiPareto model\raisebox{1.5pt}{\scriptsize*} \\
\raisebox{1.5pt}{$\centerdot$} start time relative to previous flow in session sampled from fitted Lognormal model\raisebox{1.5pt}{\scriptsize*} \\

@@

\multirow{5}{*}[2.5pt]{\class{Packet}}
& \multirow{1}{*}[-0.05em]{\submodel{Trace}} &

@@

\raisebox{1.5pt}{$\centerdot$} packet sizes and intervals drawn from fitted Exponential distributions\raisebox{1.5pt}{\scriptsize*}

### 2008-01-21 paper.tex
Redone intro; added paragraph to related work; etc.
In the process of overhauling the Methodology section.

\newcommand{\thetitle}{LAN Workload: The Failure of Uniform Modeling}

@@

\title{
\thetitle%\\
%\vspace{0.2em}
%\LARGE{On The Inadequacy of Uniform Modeling for WLAN Workloads}
}
\author{
{\large{Stefan~Karpinski, Elizabeth~M.~Belding, Kevin~C.~Almeroth}}\\
\textit{\{sgk,ebelding,almeroth\}@cs.ucsb.edu}\vspace{0.5em}\\
Department of Computer Science\\
University of California, Santa Barbara
\vspace{-0.5em}
}

\newcommand{\Section}[1]{\S\ref{sec:#1}}

@@

More than twenty years of research in analysis and modeling of network traffic have yielded tremendous advances in our understanding of the complex behaviors that emerge from the interaction of millions of humans and billions of computers on the Internet and in local area networks~(\caps{LAN}s). The rate of packet arrivals at routers on the Internet backbone is known to exhibit complex, self-similar patterns with long-range dependence~\cite{Paxson95,Paxson96,Willinger98}. Other research~\cite{Hernandez06:dissertation} has demonstrated how to accurately model the behavior of a single flow of packets between two application endpoints, including ``closed-loop'' feedback (see \Section{related-work}) induced by the ubiquitous transmission control protocol~(\caps{TCP}). Recently, the high-level statistical properties of wireless \caps{LAN} traffic have been rigorously modeled with a variety of parametric distributions~\cite{Hernandez06:wlan-traffic}.

Despite great breakthroughs in understanding and methodology, the networking community remains without a cookbook for generating realistic \caps{LAN} workloads. Worse still, it is unclear what ``realistic'' should mean in this context. Foremost, a clear and testable definition of realism for traffic models is needed. In the absence of such a definition, it remains standard in wireless experiments to generate traffic using the simplest possible model: all nodes, flows and packets are assumed to behave uniformly. Flows are commonly modeled as constant bit-rate (\caps{CBR}) streams of identical packets sent at an invariant rate; the flows in an experiment are commonly given the same duration and data rate; nodes in the network are often selected as flow endpoints randomly or arbitrarily. While more complex approaches have been proposed~\cite{xxx}, the simplest, na\"ive models remain standard because no objective criterion exists by which to compellingly argue that other models are any better. It is simply unclear that simplistic models do any harm, or how, if at all, complex models provide benefit.% in exchange for their increased complexity.

%In this paper, we propose objective criteria by which to judge whether

The experimental evaluation of new wireless technologies inevitably requires network traffic. The most commonly used models for generating wireless local-area network (\caps{WLAN}) workload are simple, treating both nodes and flows uniformly, and are generally understood to be na\"ive. They remain popular because of the lack of compelling alternatives. The scope of impact of different workloads on network performance is largely unresearched and thus, poorly understood. We believe that one of the major handicaps preventing wireless experiments from more accurately predicting real-world performance is the lack of realistic models of \caps{WLAN} workload.

@@

One of the hardest problems in modeling phenomena as complex as network traffic patterns lies in determining which characteristics of any given exemplar of behavior are essential and which can safely be abstracted away. In \caps{WLAN} traffic modeling, important characteristics have largely been determined based on intuition. In previous work~\cite{Karpinski07:realism}, we introduced a fundamentally different approach to evaluating the realism of workload models. Rather than subjectively choosing statistical measures that may or may not actually influence network performance, the realism of a model is measured directly in terms of its ability to accurately reproduce critical performance metrics. We define a traffic model to be \textit{sufficiently realistic} with respect to a given performance metric, if the model produces metric values that are similar to those observed using the original trace to generate traffic. Using this new approach, we explore the space of traffic models, establishing rigorously for the first time that common but simplistic models of \caps{WLAN} traffic drastically misrepresent important performance metrics at all levels of the protocol stack.

\begin{table}[t]
\vspace{0.7em}
\begin{tabular}[t]{|l|}
\hline
\begin{minipage}[l]{3.315in}
\begin{small}
\vspace{0.6em}
\input{properties.tex}
\vspace{0.6em}
\end{small}
\end{minipage}
\\\hline
\end{tabular}
\caption{Summary of classes and their attributes in the proposed Traffic Object Model ({\footnotesize{TOM}}) for {\footnotesize{WLAN}} workload generation.}
\vspace{-2em}
\end{table}

\begin{table*}[t]
\begin{center}
\small
\input{models.tex}
\footnotesize\\[0.5em]
\raisebox{1pt}{*}Parametric models fitted to inherited empirical distributions using maximum likelihood estimation ({\scriptsize{MLE}}).
\end{center}
\vspace{-0.75em}
\caption{Summary of how the models at each behavioral level override or inherit properties in the Traffic Object Model ({\footnotesize{TOM}}).}
\vspace{-1.75em}
\end{table*}

Our analysis shows that the standard constant bit-rate (\caps{CBR}) traffic model, with uniformly chosen nodes and uniform flows of traffic, significantly misrepresents important performance metrics at every level of the network. To qualify these results in more depth, we also investigate partially synthetic variants of this standard model, allowing us to reason about which aspects of workload models have the most impact. For the most commonly used model, the delay of application data traversing the network from sender to receiver is underestimated by more than a factor of 5, \textit{on average}, and by more than a factor of 11 in a quarter of usage scenarios. Network control overhead is overestimated, on average, by a factor of about 2.5, while link control overhead is underestimated by almost a factor of 2. Even for metrics where the average and median misrepresentations are not so extreme, the statistical characteristics of error values often indicate that something is fundamentally unrealistic. None of the partially synthetic models manage to accurately represent more than one or two performance characteristics. We further use the \caps{AODV}~\cite{rfc:aodv} and \caps{OLSR}~\cite{rfc:olsr} ad~hoc routing protocols to show that the relative performance of protocols can be inverted when changing from using a simplistic traffic model to using real traffic: using random end-point, uniform \caps{CBR} traffic, \caps{AODV} appears to induce less link-layer overhead than \caps{OLSR}, whereas using real traffic, it in fact induces more.

The first major contribution of this work is the definition of ``sufficient realism,'' together with the analytical and statistical methodology to rigorously test whether synthetic traffic models meet this definition. The second contribution is the conclusion that the most commonly used model of wireless traffic drastically and consistently misrepresents some of the most important metrics for performance evaluation of wireless protocols. Many performance comparisons based on this model may need revisiting. The last and most important contribution is the collection of ideas and analytical tools necessary to create more realistic synthetic traffic models in the future. By applying the methods developed in this work, it will be possible to discover precisely which aspects of network usage affect the realism of performance results. Once this is known, it will become feasible to create models that accurately reproduce those aspects of real wireless usage.

@@

@@

\vspace{0.5em}
\begin{samepage}
\noindent\textbf{Definition.}~A model of user behavior is \textit{sufficiently realistic} if, when compared with actual user behavior, the model, with parameter values extracted from the real data, yields statistically equivalent performance results.
\end{samepage}
\vspace{0.5em}

@@

\begin{figure}
\begin{center}
\includegraphics[width=3.5in]{nodes-flows}%
\vspace{-0.75em}%
\caption{The number of active nodes and flows over time.} 
\label{fig:nodes-flows}
\end{center}
\vspace{-2em}
\end{figure}

Our general methodology is to compare performance metrics in simulations using real traffic patterns from traces to the same metrics in simulations using a variety of synthetic traffic models, including the standard random, uniform \caps{CBR} model. For our analysis, we use a 24-hour trace recorded in an infrastructured 802.11g wireless \caps{LAN} with 18 access points, deployed at the 60th Internet Engineering Task Force meeting (\caps{IETF60}), held in San Diego during August of 2004. The traffic trace was captured using \texttt{\small{tcpdump}} at a single router, through which all wireless traffic for the meeting was routed, including traffic between wireless nodes. The snap length of the capture was 100 bytes, allowing \caps{IP}, \caps{ICMP}, \caps{UDP} and \caps{TCP} headers to be analyzed. We limit our work to the 24-hour sub-trace recorded on Wednesday, August 4th. This trace contains a broad variety of behaviors and entails a very large volume of traffic: 2.1 million flows, 58 million packets, and 52 billion bytes.
% 2,085,563 flows, 58,377,994 packets of application data, and 51,988,245,537 bytes of application data.

We do not assume or claim that the traffic found at \caps{IETF60} is representative of conference settings in general. The observed behaviors are also unlikely to resemble those found in a typical commercial or residential setting. We have chosen this trace, however, because within it can be found behaviors resembling many different types of wireless usage cases. Figure~\ref{fig:nodes-flows} shows the wide variations in the number of active flows and nodes over the course of the trace. In the night and morning hours, the traffic patterns are similar to those one might find in a moderately trafficked business or residential area. During working group sessions, we see highly concentrated, heavy usage patterns. At the zenith of activity, over 800 users, 33 thousand flows, and 1 million packets are seen in a single 10-minute trace segment. At the nadir, a lone node sent only a single 61-byte packet over the course of 10 minutes. All levels of activity between these extremes are represented. Moreover, the mix of traffic types observed changes dramatically over the course of the day, providing a wide representation of possible blends of behavior. This heterogeneity and extreme range of behaviors makes the \caps{IETF} data set ideal for this evaluation. The variety of activity gives us greater confidence that success or failure of traffic models is not tied to any specific network condition, but is broadly and generally applicable. %For a traffic model to be truly realistic, it must be realistic across many different usage cases.

@@

The 24-hour trace is split into 144 10-minute segments, each of which serves as the basis for a set of simulations using different traffic models. The traffic models range from a completely realistic trace-driven model, to a standard \caps{CBR} traffic model. Various partially synthetic intermediate models, described in \Section{models-metrics}, are simulated to study the impact of different aspects of traffic behavior on network performance. To preserve the fairness of the performance comparison, we keep as many features as possible constant across different traffic models. The traffic generated by each synthetic model preserves as many characteristics from the original trace as possible, within the constraints of the model. Moreover, the following features are preserved across all models: the numbers of wireless nodes, the number of flows, the number of application-initiated data units sent, the total bytes of application data sent, and the average flow duration (and therefore the average data rate).

\subsection{Traffic Models \& Performance Metrics}\label{sec:models-metrics}

\begin{figure*}
\begin{center}
\begin{minipage}[c]{0.67\linewidth}
\subfloat[Flow Topology: Trace]
{\includegraphics[width=1.55in]{flow_topology_trace}\label{fig:flow-topology-trace}}
\subfloat[Flow Topology: Sink]
{\includegraphics[width=1.55in]{flow_topology_sink}\label{fig:flow-topology-sink}}
\subfloat[Flow Topology: Uniform]
{\includegraphics[width=1.55in]{flow_topology_uniform}\label{fig:flow-topology-uniform}}
\subfloat[Flow Behavior: Trace]
{\includegraphics[width=1.55in]{flow_behavior_trace}\label{fig:flow-behavior-trace}}
\subfloat[Flow Behavior: Uniform]
{\includegraphics[width=1.55in]{flow_behavior_uniform}\label{fig:flow-behavior-uniform}}
\subfloat[Packet Behavior]
{\includegraphics[width=1.55in]{flow_packet_behavior}\label{fig:packet-behavior}}
\end{minipage}\hfill
\begin{minipage}[c]{0.30\linewidth}
\caption{Examples illustrating the different traffic models for the three levels of behavior. Figures \ref{fig:flow-topology-trace}, \ref{fig:flow-topology-sink} and \ref{fig:flow-topology-uniform}, show example flow topologies. The width of each line is proportional to the logarithm of the number of flows between the nodes it connects. In each graph, node zero is the gateway to the Internet. Uniform and trace flow behavior examples are plotted in Figures \ref{fig:flow-behavior-trace} and \ref{fig:flow-behavior-uniform}. The time axis indicates when the various flows start and end; the width of each flow line is proportional to the logarithm of its data rate. Flow numbers are assigned arbitrarily. Figure \ref{fig:packet-behavior} compares packet behavior for the uniform model (i.e. {\footnotesize{CBR}}), with the trace of an actual flow. In the uniform model, the cumulative data sent increases smoothly over time (gray diagonal line). In the actual packet trace, the data transmissions are variable both in size and in inter-transmission interval, leading to a ``lumpy'' cumulative data history (black step-function).}
\label{fig:traffic-models}
\end{minipage}
\end{center}
\vspace{-1em}
\end{figure*}

@@

@@

@@

@@

The simulations described in \Section{models-metrics} provide us with the raw data to compare performance metrics for synthetic traffic models with those for real traffic traces. To assess the realism of these models, however, we need a measure of how inaccurate the synthetic performance values are when compared to the real values. Let $x$ be the value of a performance metric using real traffic, and $y$ the value of the same metric using an alternate traffic model, $\M$. Some common measures of error are the difference [$y-x$], the ratio [$y/x$], and the standard error [$(y-x)/x$]. These are all reasonable measures of error; but which is most appropriate for assessing the realism of performance metrics? Instead of picking one arbitrarily, we will first consider the properties that an ideal error function should have, and then use those properties to determine the best measure of error. Moreover, we will show that the unique measure of error that exhibits these ideal properties is the  \textit{log-ratio} of metric values:
\begin{align}
\log(y/x)=\log(y)-\log(x).
\end{align}
In this discussion, $E(x,y)$ is a generic error function applied to the synthetic value, $y$, with respect to the real value, $x$.

@@

The most significant limitation on the generality of this analysis is that it is based entirely on a single data set from \caps{IETF60}---albeit a large and varied one. It is possible that traffic in this trace happens to produce network performance that is unusually dissimilar to standard traffic models. This data set, however, represents a highly heterogeneous collection of network usage behaviors, from slow and steady off-peak usage, to extremely heavy peak usage: over 800 users, 33 thousand flows, and 1 million packets in a single 10-minute trace segment. Despite the broad variety of behaviors, the results are consistent: in all types of usage scenarios, simplistic traffic models, like uniform \caps{CBR}, systematically skew important performance measurements at all levels of the network. While the precise results for other data sets might differ, it is very unlikely that \caps{CBR} traffic models will happen to accurately reproduce realistic performance in other experiments. This paper provides strong evidence that better traffic models are needed for performance evaluations.

>>>>

\newcommand{\thetitle}{On the Inadequacy of Uniform Modeling for WLAN Workload Generation}

@@

\title{\thetitle}
\author{
{\large{Stefan~Karpinski, Elizabeth~M.~Belding, Kevin~C.~Almeroth}}\\
\textit{\{sgk,ebelding,almeroth\}@cs.ucsb.edu}\vspace{0.5em}\\
Department of Computer Science\\
University of California, Santa Barbara
\vspace{-0.5em}
}

\newcommand{\Section}[1]{\hyperref[sec:#1]{Section~\ref*{sec:#1}}}
\newcommand{\FHC}{Hern\'andez-Campos~\textit{et~al.}}
\newcommand{\class}[1]{\textsf{\small{#1}}}
\newcommand{\submodel}[1]{\textsf{\small{#1}}}
\newcommand{\model}[1]{#1}
\newcommand{\RFC}[1]{\caps{RFC}~{#1}}

@@

More than twenty years of research in analysis and modeling of network traffic have yielded tremendous advances in our understanding of the complex behaviors that emerge from the interaction of millions of humans and billions of computers on the Internet and in local area networks~(\caps{LAN}s). The rate of packet arrivals at routers on the Internet backbone is known to exhibit complex, self-similar patterns with long-range dependence~\cite{Paxson95,Paxson96,Willinger98}. Other research~\cite{Hernandez06:dissertation} has demonstrated how to accurately model the behavior of a single flow of packets between two application endpoints, including ``closed-loop'' feedback induced by the ubiquitous transmission control protocol~(\caps{TCP}). Recently, the high-level statistical properties of wireless \caps{LAN} traffic have been rigorously modeled with a variety of parametric distributions~\cite{Hernandez06:wlan-traffic}.

Despite great breakthroughs in understanding and methodology, the networking community remains without a cookbook for generating realistic \caps{LAN} workloads. Worse still, it is unclear what ``realistic'' should mean in this context. Foremost, a clear and testable definition of realism for traffic models is needed. In the absence of such a definition, it remains standard in wireless experiments to generate traffic using the simplest possible model: all nodes, flows and packets are assumed to behave uniformly. Flows are commonly modeled as constant bit-rate (\caps{CBR}) streams of identical packets sent at an invariant rate; the flows in an experiment are commonly given the same duration and data rate; nodes in the network are often selected as flow endpoints randomly or arbitrarily. While more complex approaches have been proposed~\cite{Sommers04,Avallone04,Avallone06}, % TODO: better references here.
the simplest, most na\"ive models remain standard because no objective criterion exists by which to compellingly argue that other models are any better. It is simply unclear that simplistic models do any harm, or how, if at all, complex models provide benefit.

To address this foundational problem in network traffic modeling, we have introduced a fundamentally different approach to evaluating the realism of workload models~\cite{Karpinski07:realism,Karpinski07:cbr-failure}. Rather than subjectively choosing statistical measures that may or may not actually influence network performance, the realism of a model is measured directly in terms of its ability to accurately reproduce critical performance metrics. We define a traffic model to be \textit{sufficiently realistic} with respect to a given performance metric, if the model produces metric values similar to those observed using the original trace to generate traffic. Using this new approach, we explore the space of traffic models, establishing rigorously that the commonly used uniform models for node, flow and packet behavior drastically misrepresent important \caps{WLAN} performance metrics at all levels of the protocol stack.

It is natural to compare these common, na\"ive models with the high-level parametric models proposed by \FHC~\cite{Hernandez06:wlan-traffic}. Here, however, we face another fundamental challenge in traffic modeling. Even with specific models for various aspects of network behavior, it remains non-trivial to combine them into a coherent composite model capable of generating concrete workloads. The {\FHC} models, for example, specify the distribution of flows per node in the network; since every flow has two end-points, however, it remains unclear how to assign end-points to flows such that the marginal distribution of flow per node matches the BiPareto model indicated. Similarly, the models dictate BiPareto marginal distribution of flow sizes across the entire network, but say nothing about how these sizes are to be distributed among the nodes. We could assume that the number of flows per node is independent from the size of the flows at that node, but, as we shall see, this assumption is invalid. Moreover, the details of flow behavior---packet sizes and inter-packet intervals---remain entirely undefined.

Our point is not to denigrate the models proposed in~\cite{Hernandez06:wlan-traffic} in any way. On the contrary: these parametric models are the state of the art in modeling traffic patterns for wireless local area networks. The purpose, rather, is to point out how large a gap still remains between having accurate high-level models of marginal aspects of network behavior and being able to generate usable workloads with those models. Moreover, a great deal of ambiguity lies in that gap since there is no single obvious way to combine models.

To clarify, disambiguate and standardize the specification of generative local area traffic models, we propose the \textit{Traffic Object Model} (\textit{\caps{TOM}}). The \caps{TOM} entails a hierarchy of abstract classes: \class{Node}, \class{Flow}, and \class{Packet} are the primary classes; the \class{Network} class collects everything. Each abstract class defines an interface of properties. A traffic model defines concrete classes at each level that provide the required interfaces; each class defines its own properties by delegation, inheritance, or overriding properties of other classes. Because of the separation of the \caps{TOM} into well-defined levels, each concrete class naturally defines a sub-model, specific to that level's behavior; sub-models can be mixed and matched combinatorially.

We define a range of sub-models at each level of the \caps{TOM}, with decreasing similarity to the original trace behavior: \submodel{Trace} (reproduce original behavior), \submodel{Nonparametric}, \submodel{Parametric}, and \submodel{Uniform}. The control model uses the Trace sub-model at every level, exactly reproducing the behavior recorded from the original network. The standard na\"ive model uses the \submodel{Uniform} sub-model at all levels. The models proposed by {\FHC} are implemented using the \submodel{Parametric} sub-model at every level. The \submodel{Nonparametric} sub-models provide a measure of how much error is introduced by parametric approximation and parameter estimation.

%: objects and their properties together with rules for delegating, inheriting, and overriding the properties across the objects in the \caps{TOM} hierarchy. The \caps{TOM} allows

%The {\FHC} models, like most network behavior models, are marginal, meaning that they are fitted to data aggregated of over all nodes in the network, and not fitted on a per-node basis. With marginal models, since all information about interactions between the models is lost, so we must apply each model independently and hope that the result is good enough.

%Before we can define traffic models, however, we face the next major fundamental challenge in traffic modeling.

%In this paper, we address this fundamental problem in network traffic modeling by presenting an 
%In this paper, we propose objective criteria by which to judge whether

%The experimental evaluation of new wireless technologies inevitably requires network traffic. The most commonly used models for generating wireless local-area network (\caps{WLAN}) workload are simple, treating both nodes and flows uniformly, and are generally understood to be na\"ive. They remain popular because of the lack of compelling alternatives. The scope of impact of different workloads on network performance is largely unresearched and thus, poorly understood. We believe that one of the major handicaps preventing wireless experiments from more accurately predicting real-world performance is the lack of realistic models of \caps{WLAN} workload.

@@

%One of the hardest problems in modeling phenomena as complex as network traffic patterns lies in determining which characteristics of any given exemplar of behavior are essential and which can safely be abstracted away. In \caps{WLAN} traffic modeling, important characteristics have largely been determined based on intuition. In previous work~\cite{Karpinski07:realism}, we introduced a fundamentally different approach to evaluating the realism of workload models. Rather than subjectively choosing statistical measures that may or may not actually influence network performance, the realism of a model is measured directly in terms of its ability to accurately reproduce critical performance metrics. We define a traffic model to be \textit{sufficiently realistic} with respect to a given performance metric, if the model produces metric values that are similar to those observed using the original trace to generate traffic. Using this new approach, we explore the space of traffic models, establishing rigorously for the first time that common but simplistic models of \caps{WLAN} traffic drastically misrepresent important performance metrics at all levels of the protocol stack.

%Our analysis shows that the standard constant bit-rate (\caps{CBR}) traffic model, with uniformly chosen nodes and uniform flows of traffic, significantly misrepresents important performance metrics at every level of the network. To qualify these results in more depth, we also investigate partially synthetic variants of this standard model, allowing us to reason about which aspects of workload models have the most impact. For the most commonly used model, the delay of application data traversing the network from sender to receiver is underestimated by more than a factor of 5, \textit{on average}, and by more than a factor of 11 in a quarter of usage scenarios. Network control overhead is overestimated, on average, by a factor of about 2.5, while link control overhead is underestimated by almost a factor of 2. Even for metrics where the average and median misrepresentations are not so extreme, the statistical characteristics of error values often indicate that something is fundamentally unrealistic. None of the partially synthetic models manage to accurately represent more than one or two performance characteristics. We further use the \caps{AODV}~\cite{rfc:aodv} and \caps{OLSR}~\cite{rfc:olsr} ad~hoc routing protocols to show that the relative performance of protocols can be inverted when changing from using a simplistic traffic model to using real traffic: using random end-point, uniform \caps{CBR} traffic, \caps{AODV} appears to induce less link-layer overhead than \caps{OLSR}, whereas using real traffic, it in fact induces more.

%The first major contribution of this work is the definition of ``sufficient realism,'' together with the analytical and statistical methodology to rigorously test whether synthetic traffic models meet this definition. The second contribution is the conclusion that the most commonly used model of wireless traffic drastically and consistently misrepresents some of the most important metrics for performance evaluation of wireless protocols. Many performance comparisons based on this model may need revisiting. The last and most important contribution is the collection of ideas and analytical tools necessary to create more realistic synthetic traffic models in the future. By applying the methods developed in this work, it will be possible to discover precisely which aspects of network usage affect the realism of performance results. Once this is known, it will become feasible to create models that accurately reproduce those aspects of real wireless usage.

@@

This work extends that published in~\cite{Karpinski07:realism}~and~\cite{Karpinski07:cbr-failure}, using the same methodology for defining and quantifying the realism of synthetic traffic models. The Traffic Object Model builds on the combinatorial approach taken in~\cite{Karpinski07:cbr-failure}, while providing an flexible and general object-oriented framework for constructing generative traffic models. The traffic models evaluated here overlap and augment those evaluated in both prior works; comparison with the models proposed in {\FHC} is entirely new. The Qualnet simulation methodology has been enhanced by the implementation of full-duplex, trace-driven raw \caps{IP}, \caps{TCP}, and \caps{UDP} flows, and performance metric collection based on the \caps{IETF} recommendations for \caps{IP} performance metrics framework~\cite{rfc:ip-metrics,rfc:jitter}.

@@

\vspace{0.5em}
\begin{samepage}
\noindent\textbf{Definition.}~A model of user behavior is \textit{sufficiently realistic} if, compared with actual user behavior, the model, with parameter values extracted from the real data, yields statistically equivalent performance results.
\end{samepage}
\vspace{0.5em}

@@

\begin{figure}
\vspace{0.45em}
\begin{center}
\includegraphics[width=3.5in]{nodes-flows}%
\vspace{-0.75em}%
\caption{The number of active nodes and flows over time.} 
\label{fig:nodes-flows}
\end{center}
\vspace{-2em}
\end{figure}

Our general methodology is to compare performance metrics in simulations using real traffic patterns from traces to the same metrics in simulations using a variety of synthetic traffic models defined using the Traffic Object Model (\caps{TOM}) framework. We use a 24-hour trace recorded in an infrastructured 802.11g wireless \caps{LAN} with 18 access points, deployed at the 60th Internet Engineering Task Force meeting (\caps{IETF}60), held in San Diego during August of 2004. The traffic trace was captured using \texttt{\small{tcpdump}} at a single router, through which all wireless traffic for the meeting was routed, including traffic between wireless nodes. The snap length of the capture was 100 bytes, allowing \caps{IP}, \caps{ICMP}, \caps{UDP} and \caps{TCP} headers to be analyzed. We limit our work to the 24-hour sub-trace recorded on Wednesday, August 4th. This trace contains a broad variety of behaviors and entails a very large volume of traffic: 2.1 million flows, 58 million packets, and 52 billion bytes.
% 2,085,563 flows, 58,377,994 packets of application data, and 51,988,245,537 bytes of application data.

We do not assume or claim that the traffic found at \caps{IETF}60 is representative of conference settings in general. The observed behaviors are also unlikely to resemble those found in a typical commercial or residential setting. We have chosen this trace, however, because within it can be found behaviors resembling many different types of wireless usage cases. Figure~\ref{fig:nodes-flows} shows the wide variations in the number of active flows and nodes over the course of the trace. In the night and morning hours, the traffic patterns are similar to those one might find in a moderately trafficked business or residential area. During working group sessions, we see highly concentrated, heavy usage patterns. At the zenith of activity, over 800 users, 33 thousand flows, and 1 million packets are seen in a single 10-minute trace segment. At the nadir, a lone node sent only a single 61-byte packet over the course of 10 minutes. All levels of activity between these extremes are represented. Moreover, the mix of traffic types observed changes dramatically over the course of the day, providing a wide representation of possible blends of behavior. This heterogeneity and extreme range of behaviors makes the \caps{IETF} data set ideal for this evaluation. The variety of activity gives us greater confidence that success or failure of traffic models is not tied to any specific network condition, but is broadly and generally applicable. %For a traffic model to be truly realistic, it must be realistic across many different usage cases.

@@

\begin{table}[t]
\vspace{0.75em}
\begin{center}
\begin{tabular}[t]{|l|}
\hline
\begin{minipage}[l]{2.8in}
\begin{small}
\vspace{0.6em}
\input{properties.tex}
\vspace{0.6em}
\end{small}
\end{minipage}
\\\hline
\end{tabular}
\end{center}
\vspace{-0.5em}
\caption{Classes and their properties in the Traffic Object Model.}
\vspace{-2em}
\end{table}

\subsection{The Traffic Object Model~(\caps{TOM})}\label{sec:models}

\begin{table*}[t]
\begin{center}
\small
\input{models.tex}
\footnotesize\\[0.5em]
\raisebox{1pt}{*}Parametric models fitted to inherited empirical distributions using maximum likelihood estimation ({\scriptsize{MLE}}).
\end{center}
\vspace{-0.75em}
\caption{Summary of how the sub-models at each behavioral level of the {\footnotesize{TOM}} delegate, override or inherit properties.}
\vspace{-1.75em}
\end{table*}

@@

\begin{figure*}[tb]
\begin{center}
\subfloat[trace node behavior]
{\includegraphics[width=1.392in]{flow_topology_trace}\label{fig:flow-topology-trace}}
\subfloat[uniform node behavior]
{\includegraphics[width=1.392in]{flow_topology_uniform}\label{fig:flow-topology-uniform}}
\subfloat[trace flow behavior]
{\includegraphics[width=1.392in]{flow_behavior_trace}\label{fig:flow-behavior-trace}}
\subfloat[uniform flow behavior]
{\includegraphics[width=1.392in]{flow_behavior_uniform}\label{fig:flow-behavior-uniform}}
\subfloat[packet behavior]
{\includegraphics[width=1.392in]{flow_packet_behavior}\label{fig:packet-behavior}}
\caption{Examples illustrating trace versus uniform behaviors a the node, flow and packet levels. Figures \ref{fig:flow-topology-trace} and \ref{fig:flow-topology-uniform} show example node behaviors. The width of each line is proportional to the logarithm of the number of flows between the nodes it connects (zero is the Internet gateway). Uniform and trace flow behavior examples are plotted in Figures \ref{fig:flow-behavior-trace} and \ref{fig:flow-behavior-uniform}. The time axis indicates when flows start and end; the width of each flow line is proportional to the logarithm of its data rate. Figure \ref{fig:packet-behavior} compares \caps{CBR} packet behavior with the trace of an actual flow. In the uniform model, the cumulative data sent increases smoothly over time, whereas in the actual packet trace, the transmissions are variable both in size and in inter-packet interval, leading to a ``lumpy'' cumulative data plot.}
\label{fig:traffic-models}
\end{center}
\vspace{-2em}
\end{figure*}

@@

\subsection{Performance Metrics}\label{sec:metrics}

@@

The 24-hour trace is split into 144 10-minute segments, each of which serves as the basis for a set of simulations using different traffic models. The traffic models range from a completely realistic trace-driven model, to a standard \caps{CBR} traffic model. Various partially synthetic intermediate models, described in \Section{models}, are simulated to study the impact of different aspects of traffic behavior on network performance. To preserve the fairness of the performance comparison, we keep as many features as possible constant across different traffic models. The traffic generated by each synthetic model preserves as many characteristics from the original trace as possible, within the constraints of the model. Moreover, the following features are preserved across all models: the numbers of wireless nodes, the number of flows, the number of application-initiated data units sent, the total bytes of application data sent, and the average flow duration (and therefore the average data rate).

@@

The simulations described in \Section{simulations} provide us with the raw data to compare performance metrics for synthetic traffic models with those for real traffic traces. To assess the realism of these models, however, we need a measure of how inaccurate the synthetic performance values are when compared to the real values. Let $x$ be the value of a performance metric using real traffic, and $y$ the value of the same metric using an alternate traffic model, $\M$. Some common measures of error are the difference [$y-x$], the ratio [$y/x$], and the standard error [$(y-x)/x$]. These are all reasonable measures of error; but which is most appropriate for assessing the realism of performance metrics? Instead of picking one arbitrarily, we will first consider the properties that an ideal error function should have, and then use those properties to determine the best measure of error. Moreover, we will show that the unique measure of error that exhibits these ideal properties is the  \textit{log-ratio} of metric values:
\begin{align}
\log(y/x)=\log(y)-\log(x).
\end{align}
In this discussion, $E(x,y)$ is a generic error function applied to the synthetic value, $y$, with respect to the real value, $x$.

@@

The most significant limitation on the generality of this analysis is that it is based entirely on a single data set from \caps{IETF}60---albeit a large and varied one. It is possible that traffic in this trace happens to produce network performance that is unusually dissimilar to standard traffic models. This data set, however, represents a highly heterogeneous collection of network usage behaviors, from slow and steady off-peak usage, to extremely heavy peak usage: over 800 users, 33 thousand flows, and 1 million packets in a single 10-minute trace segment. Despite the broad variety of behaviors, the results are consistent: in all types of usage scenarios, simplistic traffic models, like uniform \caps{CBR}, systematically skew important performance measurements at all levels of the network. While the precise results for other data sets might differ, it is very unlikely that \caps{CBR} traffic models will happen to accurately reproduce realistic performance in other experiments. This paper provides strong evidence that better traffic models are needed for performance evaluations.
